EPA PM2.5 emissions
========================================================
Author: Sheng Li

**Q1.** Have total emissions from PM2.5 decreased in the United States from 1999 to 2008? Using the baseplotting system, make a plot showing the total PM2.5 emission from all sources for each of the years 1999, 2002, 2005, and 2008.

```{r}
# Load the data
NEI <- readRDS("summarySCC_PM25.rds")
SCC <- readRDS("Source_Classification_Code.rds")

# Aggregate total emissions from PM2.5, convert from tons to kilotons
attach(NEI)
EmissionsPM2.5 <- tapply((Emissions)/1000,year,sum)
```

```{r, fig.height=6, fig.width=7}
barplot(EmissionsPM2.5, xlab = "Year", ylab = expression("Emissions of PM"[2.5]* " (in kilotons)"), 
    main = expression("Total emissions from PM"[2.5]* " in US (1999-2008)"), col = c("red", "blue", "grey", "orange") )
```

Alternative line plot:
```{r, fig.height=6, fig.width=7}
# Aggregate total emissions from PM2.5, convert from tons to kilotons
EmissionsPM2.5 <- aggregate(NEI[c("Emissions")]/1000, list(year = NEI$year), sum)

# Create the line plot
plot(EmissionsPM2.5$year, EmissionsPM2.5$Emissions, type="b", xlab = "Year", 
    ylab = expression("Emissions of PM"[2.5]* " (in kilotons)"), 
    main = expression("Total emissions from PM"[2.5]* " in US (1999-2008)"), col = "Blue")
```

**Q2.** Have total emissions from PM2.5 decreased in the Baltimore City, Maryland (fips == "24510") from 1999 to 2008? Use the base plotting system to make a plot answering this question.

```{r}
# Aggregate total emissions in Baltimore, convert from tons to kilotons
Baltimore <- subset(NEI, fips == "24510")
BaltimorePM2.5 <- tapply((Baltimore$Emissions)/1000, Baltimore$year, sum)
```

```{r, fig.height=6, fig.width=7}
# Plot the bar chart
barplot(BaltimorePM2.5, xlab = "Year", ylab = expression("Emissions of PM"[2.5]* " (in kilotons)"), 
    main = expression("Total emissions from PM"[2.5]* " in Baltimore City, MD (1999-2008)"), 
    col = c("red", "blue", "grey", "orange"))
```

Alternative line plot:
```{r, fig.height=6, fig.width=7}
# Aggregate total emissions in Baltimore, convert from tons to kilotons
Baltimore <- subset(NEI, fips == "24510")
BaltimorePM2.5 <- aggregate(Baltimore[c("Emissions")]/1000, list(year = Baltimore$year), sum)

# Create the line plot
plot(BaltimorePM2.5$year, BaltimorePM2.5$Emissions, type="b", xlab = "Year", 
    ylab = expression("Emissions of PM"[2.5]* " (in kilotons)"), 
    main = expression("Total emissions from PM"[2.5]* " in Baltimore City, MD (1999-2008)"), col = "Blue")
```

**Q3.** Of the four types of sources indicated by the type (point, nonpoint, onroad, nonroad) variable, which of these four sources have seen decreases in emissions from 1999–2008 for Baltimore City? Which have seen increases in emissions from 1999–2008? Use the ggplot2 plotting system to make a plot answer this question.

```{r}
require(ggplot2)
Baltimore <- subset(NEI, fips == "24510")
BaltimorePM2.5 <- aggregate(Baltimore[c("Emissions")], list(type = Baltimore$type, year = Baltimore$year), sum)
```

```{r, fig.height=6, fig.width=7}
# Create line plot
qplot(year, Emissions, data = BaltimorePM2.5, color = type, geom = "path", xlab = "Year", 
    ylab = expression("Emissions of PM"[2.5]* " (in tons)"), 
    main = expression("Total emissions from PM"[2.5]* " in Baltimore City, MD (1999-2008)"))
```

**Q4.** Across the United States, how have emissions from coal combustion-related sources changed from 1999–2008?

```{r}
# Merge coal related SCC and NEI data set
SCC.coal <- SCC[grepl("coal", SCC$Short.Name, ignore.case=TRUE),]
CoalEmission <- merge(NEI, SCC.coal, by = "SCC")
# Aggregate total emissions from coal-related combustion, convert from tons to kilotons
CoalEmissionPM2.5 <- aggregate(CoalEmission[, "Emissions"]/1000, by = list(CoalEmission$year), sum)
colnames(CoalEmissionPM2.5) <- c("year", "Emissions")
```

```{r, fig.height=6, fig.width=7}
# Create line plot
ggplot(CoalEmissionPM2.5, aes(x = year, y = Emissions)) + 
  geom_line(aes(group = 1, col = Emissions)) + geom_point(aes(size = 1, col = Emissions)) + 
  ylab(expression("Emissions of PM"[2.5]* " (in kilotons)")) + xlab("Year") +
  ggtitle("Total emissions from Coal Combustion (1999-2008)")
```

**Q5.** How have emissions from motor vehicle sources changed from 1999–2008 in Baltimore City?

```{r}
# Aggregate total emissions from motor vehicles in Baltimore
Baltimore <- subset(NEI, fips == "24510" & type == "ON-ROAD")
Baltimore$year <- factor(Baltimore$year, levels = c('1999', '2002', '2005', '2008'))
BaltimorePM2.5 <- aggregate(Baltimore[, "Emissions"], by = list(Baltimore$year), sum)
colnames(BaltimorePM2.5) <- c("year", "Emissions")
```

```{r, fig.height=6, fig.width=7}
# Create barplot
ggplot(BaltimorePM2.5, aes(x = year, y = Emissions)) + geom_bar(aes(fill=year), stat="identity") +  
  ylab(expression("Emissions of PM"[2.5]* " (in tons)")) + xlab("Year") + 
  ggtitle("Total Emissions of Motor Vehicles in Baltimore") +
  geom_text(aes(label = round(Emissions), size = 3, hjust = 1, vjust = 0))
```

Alternative line plot:
```{r, fig.height=6, fig.width=7}
ggplot(BaltimorePM2.5, aes(x = year, y = Emissions)) + 
    geom_line(aes(group = 1, col = Emissions)) + 
    ylab(expression("Emissions of PM"[2.5]* " (in tons)")) + xlab("Year") + 
    ggtitle("Total Emissions of Motor Vehicles in Baltimore")
```

**Q6.** Compare emissions from motor vehicle sources in Baltimore City with emissions from motor vehicle sources in Los Angeles County, California (fips == "06037"). Which city has seen greater changes over time in motor vehicle emissions?

```{r}
# Aggregate total emissions from motor vehicles in Baltimore
Baltimore <- subset(NEI, fips == "24510" & type == "ON-ROAD")
Baltimore$year <- factor(Baltimore$year, levels = c('1999', '2002', '2005', '2008'))
BaltimorePM2.5 <- cbind(aggregate(Baltimore[, "Emissions"], by = list(Baltimore$year), sum), City = "Baltimore")
colnames(BaltimorePM2.5) <- c("year", "Emissions", "City")

# Aggregate total emissions from motor vehicles in Los Angeles
LosAng <- subset(NEI, fips == "06037" & type == "ON-ROAD")
LosAng$year <- factor(LosAng$year, levels = c('1999', '2002', '2005', '2008'))
LosAngPM2.5 <- cbind(aggregate(LosAng[, "Emissions"], by = list(LosAng$year), sum), City = "Los Angeles")
colnames(LosAngPM2.5) <- c("year", "Emissions", "City")

BaltLA <- rbind.data.frame (BaltimorePM2.5, LosAngPM2.5)
```

```{r, fig.height=6, fig.width=7}
# Plot bar chart for comparison 
ggplot(BaltLA, aes(year, Emissions)) + geom_bar(aes(fill = year), stat = "identity") + facet_grid(. ~ City) + 
    ggtitle("Total Emissions from motor vehicles in Baltimore and in Los Angeles") + 
    ylab(expression("Emissions of PM"[2.5]* " (in tons)")) + xlab("Year") + 
    geom_text(aes(label=round(Emissions), size = 2, hjust = 1, vjust = -1))
```

Alternative comparison in 3-year % change:
```{r}
pairs <- c(1999, 2002, 2005, 2008)
percentchange <- round(data.frame(sapply(1:3, FUN = function(x) 
          (BaltLA$Emissions[BaltLA$year == pairs[x + 1]] - BaltLA$Emissions[BaltLA$year == pairs[x]]) / BaltLA$Emissions[BaltLA$year == pairs[x]])) * 100, 1)
percentchange <- rbind(t(percentchange[1, ]),t(percentchange[2, ]))
percentchange <- as.data.frame(cbind(rep(c("1999 to 2002", "2002 to 2005", "2005 to 2008"), 2), c(rep("Baltimore", 3), rep("Los Angeles", 3)), percentchange))
colnames(percentchange) <- c("Period", "City", "Change")
percentchange$Change <- as.numeric(as.character(percentchange$Change))
```

```{r, fig.height=6, fig.width=8}
# Plot bar chart for percentage change comparison 
ggplot(percentchange, aes(Period, Change)) + geom_bar(aes(fill = City), stat = "identity") + facet_grid(. ~ City) + 
    ggtitle("% Change in Total Emissions from motor vehicles in Baltimore vs. L.A.") + 
    ylab(expression("3 Years % Change in PM"[2.5])) + xlab("Year") + 
    geom_text(aes(label=round(Change), size = 2, hjust = 1, vjust = -1))
```